A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-05-21 sections/Discussion.tex
Modification according to comments and known typos
Lexically scoped type variable extension

@@

Moreover, existential variables need special treatment, since they may be solved at any point,
or be recycled if not solved within its scope.
If an existential variable is solved, we not only propagate the solution to the other judgments,
but also replace occurrences in the elaboration list.
If an existential variable is recycled, meaning that it is not constrained,
we can pick any well-formed monotype as its solution.
The unit type $1$, as the simplest type in the system, is a good choice.

>>>>

\jimmy{TODO: update the summary}

@@

Moreover, existential variables need special treatment, since they may be solved at any point,
or be recycled if not solved within their scopes.
If an existential variable is solved, we not only propagate the solution to the other judgments,
but also replace occurrences in the elaboration list.
If an existential variable is recycled, meaning that it is not constrained,
we can pick any well-formed monotype as its solution.
The unit type $1$, as the simplest type in the system, is a good choice.

\subsection{Lexically-scoped type variables}
We further extend the type system to support a simple extension, Lexically-scoped type variables.
From a practical point of view,
such extension enables the programmer to refer to type variables in a generic function signature
when implementing of that function.
For example,
$$(\lam{\lam[y] (x:a)}) : \all[a~b] a \to b \to a$$
has an inner annotation $(x:a)$ that refers to the outer generic type variable $a$.
This is not a surprising feature, since the declarative system already accepts these kind of programs
$$
\inferrule*[right=$\mathtt{DeclAnno}$]
{
    \Psi \vdash \all A \quad
    \inferrule*[right=$\mathtt{Decl\forall I}$]
    {\Psi,a \vdash e \Lto A}
    {\Psi\vdash e\Lto \all A}
}
    {\Psi\vdash (e: \all A)\To \all A}
$$

But there are issues about the well-formedness condition.
Normally $\Gm \vdash_{wf} (e : A)$ is defined by $\Gm \vdash_{wf} e$ and $\Gm \vdash_{wf} A$.
When $A = \all A'$, the type variable $a$ is not introduced to the context of $e$,
therefore $\Gm \vdash_{wf} e$ is not derivable.

To address the problem, one can either
1) change the well-formedness judgment by introducing the binder for terms as well;
or 2) add a syntax that explicitly bind the generic type variable.
The former approach requires to resolve overlapping cases
$(e : \all A)$ with $(e : B)$ where $B \neq \all B'$.
Therefore we choose the latter approach due to its simplicity.
The syntax of expression is extended with one more constructor:
$$
\begin{array}{l@{\qquad}lcl}
\text{Expressions}\qquad&e &::=&\quad x \mid () \mid \lam e \mid e_1~e_2 \mid (e:A) \mid \Lambda a.~e:A
\end{array}
$$

Notice that the type-lambda syntax $\Lambda a.~e:A$ annotates the generic function $e$ with $\all A$,
while making $a$ visible inside the body of the function.
Corresponding rules are introduced for both the declarative system and the algorithmic system:
$$
\inferrule*[right=$\mathtt{Decl}\Lambda$]
    {\Psi,a \vdash A \quad \Psi,a \vdash e\Lto A}
    {\Psi\vdash \Lambda a.~e:A \To \all A}
$$
$$\Gm \Vdash \Lambda a.~e:A \To_b \jg \algrule \Gm \Vdash [(\all A)/b]\jg, a \Vdash e \Lto A$$

The new syntax for expressions also affect the well-formedness judgment:
$$
\inferrule*[right=$\mathtt{wf\_}\Lambda$]
    {\Gm,a \vdash_{wf} e \quad \Gm,a \vdash_{wf} A}
    {\Gm \vdash_{wf} \Lambda a.~e:A}
$$

In a practical point of view, programmers may not write the syntax $\Lambda a. e : A$ directly.
The ScopedTypeVariables extension of Haskell is effective only
when the type signature is explicitly universally quantified
(the compiler translates into an expression similar to $\Lambda a. e : A$);
otherwise the program means the normal syntax $e : \all A$,
and one may not later refer to the type variable $a$.

We also obtained all three desired properties for the extended system,
namely soundness, completeness and decidability.
